---
title: "Prediction Assignment"
author: "Felipe Augusto Azuero Mutis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r loadLibraries, message=FALSE}
library(dplyr)
library(caret)
```

```{r setSeed}
set.seed(4564)
```

## The Data

```{r loadTheData}
training <- read.csv("./data/pml-training.csv")
testing <- read.csv("./data/pml-testing.csv")
```

The database used for this study is the one gathered by [Velloso et al.](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har) In the study conducted by the authors, six healthy participants were asked to perform a set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different ways.  During the exercise, accelerometer data were collected on the belt, forearm, arm.

The data has already been divided between a training dataset and a testing dataset. Each dataset has `r ncol(training)` variables.  The training dataset has `r nrow(training)` while the testing dataset has `r nrow(testing)` observations. Only in the training dataframe we can find the column `classe` while in the testing dataframe there is a column named `problem_id`.

A summary of the training database is presented below.

```{r summaryTraining}
str(training)
```

As can be seen, several of the variables have a large number of missing values, coded in different ways. There are also some values expressed as `#DIV/0!`.

## Data Cleaning.

As a first step I will encode the values correctly.  Also due to the small number of occurrences and the large number of data I will eliminate the observations where there is one or more `#DIV/0!` present.  This operation will be carried out for both the training and testing datasets.

```{r dataCleaningFunction}

dataCleaning <- function(df) {
  out <- df %>%
    mutate(across(.fns = ~replace(., . == "", NA))) %>%
    filter(across(.fns = ~!grepl("#DIV/0!", .))) %>%
    mutate(across(.cols = c(user_name, new_window, num_window),
                  .fns = ~as.factor(.)))
  out
}

trainingClean <- dataCleaning(training)
testingClean <- dataCleaning(testing)

```

To choose the covariates, I will first eliminate those columns that have more than 50% of missing values, and other columns like the index, timestamps, and window information.

```{r dataCleaningCovariates}

trainingClean <- trainingClean %>%
  select(where(~mean(is.na(.)) < 0.5)) %>%
  select(-c(X, raw_timestamp_part_1, 
            raw_timestamp_part_2, 
            cvtd_timestamp, 
            new_window, 
            num_window))

```

## Validation

Since the database has a relatively large number of observations I will divide the training database into a part to train the models and another part to validate the results.

```{r validation}
inValidation <- createDataPartition(y=trainingClean$classe,
                                    p=0.9, list=F)
trDF <- trainingClean[inValidation,]
valDF <- trainingClean[-inValidation,]

```

## Training Models.

I will use a random forest model with 5 fold cross validation.  This decision is made because of the number of variables and observations, and to avoid overfitting. 

```{r randomForests, results = 'hide', error=F, warning=F, message=F}

modFit <- train(classe~., 
                data=trDF,
                trControl=trainControl(method="cv", number=5))

modFit

```


## References

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

